### 2024-07-30 wqo-on-words.tex
starting to write a bit

\documentclass[11pt]{article}

\input{lib/packages}
\input{lib/aliaume}
\input{src/metadata.tex}

\begin{document}
\maketitle

\input{src/introduction}
\input{src/prefixes}
\input{src/infixes}

\bibliographystyle{plainurl}
\bibliography{papers.bib}

\end{document}