### 2024-08-06 wqo-on-words.tex
rewriting introduction

\documentclass[11pt]{article}

\input{lib/packages}     % packages needed for this paper
\input{lib/aliaume}      % aliaume’s custom macrosh

\input{lib/maths}        % custom math commands specific to this paper
\input{knowledges.kl}    % textual knowledges

\input{src/metadata.tex} % metadata of the paper (generated)

\begin{document}
\maketitle

\input{src/introduction}
\input{src/prefixes}
\input{src/infixes}

\bibliographystyle{plainurl}
\bibliography{papers.bib}

\end{document}