### 2024-08-07 src/metadata.tex
writing proofs

% Description: Metadata for the plain template
%
% Warning: this file is generated automatically
% from the paper-meta.yaml file. If you really want to edit
% this file, please mirror the changes to paper-meta.yaml.

\title{Well-quasi-orderings on word languages}
\author{%
        Roland Guttenberg \and
        Nathan Lhote \and
        Aliaume Lopez \and
        Corto Mascle \and
        Vincent Michielini \and
        Lia Schütze \and
        Omid Yaghoubi}

\newcommand{\makeabstract}{
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself
    well-quasi-ordered. This seminal result by Higman is a cornerstone
    of the theory of well-quasi-orderings and has found numerous
    applications in computer science. However, this result is based on a
    specific choice of ordering on words, the (scattered) subword
    ordering. In this paper, to what extent other natural orderings
    (prefix, suffix, infix) on words can be used to derive Higman-like
    theorems. More specifically, we are interested in characterizing
    \emph{languages} of words that are well-quasi-ordered under these
    orderings. We show that a simple characterization is possible for
    the prefix and suffix orderings, and that under extra regularity
    assumptions, this also lifts to the infix ordering.
\end{abstract}
}