### 2024-10-11 wqo-on-words.tex
Add linenumbers

% !TeX document-id = {8cab1ed7-6690-4820-80f7-c18b55f758a8}
% !BIB program = bibtex
\documentclass[11pt,a4paper,twosided]{article}
\usepackage{geometry}
\usepackage{lineno}
\linenumbers
\geometry{margin=3.5cm}

\input{lib/packages}     % packages needed for this paper
\input{lib/aliaume}      % aliaume’s custom macros

\input{lib/maths}        % custom math commands specific to this paper
\input{knowledges.kl}    % textual knowledges

\input{src/metadata.tex} % metadata of the paper (generated)

\begin{document}
\maketitle
\makeabstract

% main content
\input{src/introduction}
\input{src/prefixes}
\input{src/infixes}
\input{src/conclusion}

% bibliographies
\bibliographystyle{plainurl}
\bibliography{papers.bib}

% appendices
\appendix

\end{document}