### 2025-09-08 wqo-on-words.lncs.tex
fix: ci/cd

%! TEX program = pdflatex
% WARNING: this is a generated file.
%
% Please do not edit this file directly. 
% - If you want to update the medatata of the paper (title, authors, abstract), please
%   edit the `paper-meta.yaml` file in the root of the repository.
% - If you want to update the content of the paper, please edit the latex files
%   in the `src` directory.
% - If you want to update the template itself (e.g., change the layout), please
%   edit the `templates/lncs/lncs.tex` file instead.
\documentclass[runningheads]{llncs}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{lineno}
\linenumbers

\usepackage{todonotes}

% babel for language settings
\usepackage[english]{babel}

% microtype for better typography
\usepackage{microtype}

% math packages
\usepackage{amssymb,amsmath,stmaryrd,thmtools,thm-restate,upgreek}

% graphics packages
\usepackage{graphicx}
\usepackage[obeyclassoptions,mode=tex]{standalone}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning}
\usetikzlibrary{automata}
\usetikzlibrary{tikzmark}
\usetikzlibrary{patterns}
\usetikzlibrary{arrows}
\tikzset{every state/.style={minimum size=1pt}}
\usepackage{tikz-cd}

\usepackage{subcaption}

% links inside the document
\usepackage{hyperref}
\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
\Crefname{assumption}{Assumption}{Assumptions}
\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
\knowledgeconfigure{notion}

% Tables 
\usepackage{booktabs}
\usepackage{varwidth}

% Packages for macro definitions
\usepackage{xparse}
\usepackage{xpatch}
\usepackage{tokcycle}
\usepackage{ifthen}

% Proof trees
\usepackage{bussproofs}

% Algorithms
\usepackage{algorithm2e}
\Crefname{algocfline}{Algorithm}{Algorithms}
\crefname{algocfline}{Algorithm}{Algorithms}
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

% Colors 
\usepackage{ensps-colorscheme}
\knowledgestyle{intro notion}{color={A2}, emphasize}
\knowledgestyle{notion}{color={B1}}

% we include whatever the user wants to include in the header

% we include libraries (tex files) usually written in the `lib` directory
\input{lib/aliaume}
\input{lib/maths}
\input{knowledges.kl}

\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main 0121a5929682a015f9ca5ed9399070224366a74c \\
       2025-08-12 15:10:21 +0200
}
\titlerunning{WQO on Words}

\author{
        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
\authorrunning{N. Lhote and A. Lopez and L. Schütze}
\institute{
University of Warsaw
 \and
Aix-Marseille University
 \and
Max Planck Institute for Software Systems
}

\date{\today}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered. This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science. However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering. In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems. More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings. We show that a simple characterization is possible for the prefix and suffix orderings, and that under extra regularity assumptions, this also extends to the infix ordering. We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
\end{abstract}

% Include the content of the paper
\input{src/introduction}
\input{src/prelims.tex}
\input{src/prefixes}
\input{src/infixes}
\input{src/amalgam}
\input{src/conclusion}

% Include acknowledgements

% Include the bibliography
\bibliographystyle{splncs04}
\bibliography{papers.bib}

% If there are any appendices, we include them here.

\end{document}