### 2025-09-08 wqo-on-words.lipics.tex
fix: ci/cd

%! TEX program = xelatex
% WARNING: this is a generated file.
%
% Please do not edit this file directly. 
% - If you want to update the medatata of the paper (title, authors, abstract), please
%   edit the `paper-meta.yaml` file in the root of the repository.
% - If you want to update the content of the paper, please edit the latex files
%   in the `src` directory.
% - If you want to update the template itself (e.g., change the layout), please
%   edit the `templates/lipics/lipics.tex` file instead.
\documentclass[
    a4paper,
    UKenglish,
    cleveref,
    autoref,anonymous,
    thm-restate]{lipics-v2021}

\hideLIPIcs

% graphics packages
\usepackage{graphicx}
\usepackage[obeyclassoptions,mode=tex]{standalone}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning}
\usetikzlibrary{automata}
\usetikzlibrary{tikzmark}
\usetikzlibrary{patterns}
\usetikzlibrary{arrows}
\tikzset{every state/.style={minimum size=1pt}}
\usepackage{tikz-cd}

% links inside the document
\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
\knowledgeconfigure{notion}

% Tables 
\usepackage{booktabs}
\usepackage{varwidth}

% Packages for macro definitions
\usepackage{xparse}
\usepackage{xpatch}
\usepackage{tokcycle}
\usepackage{ifthen}

% Proofs
\usepackage{bussproofs}

% Colors 
\usepackage{ensps-colorscheme}
% parametrize knowledge colors and style
% intro => A4 + emph 
% kl    => normal color, normal size not emph
\knowledgestyle{intro notion}{color={A2}, emphasize}
\knowledgestyle{notion}{color={B1}}
\hypersetup{
    colorlinks=true,
    breaklinks=true,
    linkcolor=A4,
    citecolor=A4,
    urlcolor=A4,
    filecolor=A4,
}

% we include whatever the user wants to include in the header

% we include libraries (tex files) usually written in the `lib` directory
\input{lib/aliaume}
\input{lib/maths}
\input{knowledges.kl}

% We include the title and author information based on the 
% `paper-meta.yaml` file.

\title{Well-quasi-orderings on word languages}

\titlerunning{WQO on Words}
\author{Nathan Lhote}
       {Aix-Marseille University}
       {nathan.lhote@lis-lab.fr}
       {https://orcid.org/0000-0003-3303-5368}
       {}
\author{Aliaume Lopez}
       {University of Warsaw}
       {ad.lopez@uw.edu.pl}
       {https://orcid.org/0000-0002-4205-327X}
       {}
\author{Lia Schütze}
       {Max Planck Institute for Software Systems}
       {lschuetze@mpi-sws.org}
       {https://orcid.org/0000-0003-4002-5491}
       {}

\authorrunning{N. Lhote and A. Lopez and L. Schütze}

\Copyright{Nathan Lhote and Aliaume Lopez and Lia Schütze}

\ccsdesc[100]{Mathematics of computing\textsubscript{Algebra}Polynomials}
\ccsdesc[100]{Theory of computation\textasciitilde Logic and verification\textasciitilde Model checking}

\keywords{equivariant ideal, Hilbert basis, ideal membership problem, orbit finite, oligomorphic, well-quasi-ordering} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Now, we create the document itself.
\begin{document}
% Generate the title page
\maketitle
% Print the abstract
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered. This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science. However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering. In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems. More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions. We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
\end{abstract}

% Include the content of the paper
\input{src/introduction}
\input{src/prelims.tex}
\input{src/prefixes}
\input{src/infixes}
\input{src/amalgam}
\input{src/conclusion}

% Include the bibliography
\bibliographystyle{plainurl}
\bibliography{papers.bib}

% If there are any appendices, we include them here.

\end{document}