### 2025-10-16 wqo-on-words.lncs.tex
Add thmtools integration and demonstrate restatable theorem proof of concept
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

%! TEX program = pdflatex
% WARNING: this is a generated file.
%
% Please do not edit this file directly. 
% - If you want to update the medatata of the paper (title, authors, abstract), please
%   edit the `paper-meta.yaml` file in the root of the repository.
% - If you want to update the content of the paper, please edit the latex files
%   in the `src` directory.
% - If you want to update the template itself (e.g., change the layout), please
%   edit the `templates/lncs/lncs.tex` file instead.
\documentclass[runningheads,envcountsame]{llncs}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{lineno}
\linenumbers

\usepackage{todonotes}

% babel for language settings
\usepackage[english]{babel}

% microtype for better typography
%\usepackage{microtype} % Disabled due to font expansion issues

% math packages
\usepackage{amssymb,amsmath,stmaryrd,thmtools,thm-restate,upgreek}

% Redefine theorem environments to be compatible with thmtools restatable
% llncs.cls already defines these using \spnewtheorem, but we need them to work with thmtools
% We need to undefine both the environments and the counters
\makeatletter
\let\theorem\relax
\let\endtheorem\relax
\let\c@theorem\relax
\let\thetheorem\relax
\let\p@theorem\relax
\let\lemma\relax
\let\endlemma\relax
\let\c@lemma\relax
\let\thelemma\relax
\let\p@lemma\relax
\let\corollary\relax
\let\endcorollary\relax
\let\c@corollary\relax
\let\thecorollary\relax
\let\p@corollary\relax
\let\proposition\relax
\let\endproposition\relax
\let\c@proposition\relax
\let\theproposition\relax
\let\p@proposition\relax
\let\definition\relax
\let\enddefinition\relax
\let\c@definition\relax
\let\thedefinition\relax
\let\p@definition\relax
\let\example\relax
\let\endexample\relax
\let\c@example\relax
\let\theexample\relax
\let\p@example\relax
\let\remark\relax
\let\endremark\relax
\let\c@remark\relax
\let\theremark\relax
\let\p@remark\relax
\makeatother

% Now redeclare them using thmtools for restatable support
% Using llncs-like formatting: bold for theorem names, italic for body
\declaretheorem[style=plain,numberwithin=section,name=Theorem]{theorem}
\declaretheorem[style=plain,sibling=theorem,name=Lemma]{lemma}
\declaretheorem[style=plain,sibling=theorem,name=Corollary]{corollary}
\declaretheorem[style=plain,sibling=theorem,name=Proposition]{proposition}
\declaretheorem[style=definition,sibling=theorem,name=Definition]{definition}
\declaretheorem[style=definition,sibling=theorem,name=Example]{example}
\declaretheorem[style=remark,sibling=theorem,name=Remark]{remark}

% graphics packages
\usepackage{graphicx}
\usepackage[obeyclassoptions,mode=tex]{standalone}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning}
\usetikzlibrary{automata}
\usetikzlibrary{tikzmark}
\usetikzlibrary{patterns}
\usetikzlibrary{arrows}
\tikzset{every state/.style={minimum size=1pt}}
\usepackage{tikz-cd}

\usepackage{subcaption}

% links inside the document
\usepackage{hyperref}
\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
\Crefname{assumption}{Assumption}{Assumptions}
\Crefname{conjecture}{Conjecture}{Conjectures}
\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
\knowledgeconfigure{notion}

% Tables 
\usepackage{booktabs}
\usepackage{varwidth}

% Packages for macro definitions
\usepackage{xparse}
\usepackage{xpatch}
\usepackage{tokcycle}
\usepackage{ifthen}

% Proof trees
\usepackage{bussproofs}

% Algorithms
\usepackage{algorithm2e}
\Crefname{algocfline}{Algorithm}{Algorithms}
\crefname{algocfline}{Algorithm}{Algorithms}
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

% Colors 
\usepackage{ensps-colorscheme}
\knowledgestyle{intro notion}{color={A2}, emphasize}
\knowledgestyle{notion}{color={B1}}

% we include whatever the user wants to include in the header

% we include libraries (tex files) usually written in the `lib` directory
\input{lib/aliaume}
\input{lib/maths}
\input{knowledges.kl}

\newcommand{\repositoryUrl}{Undefined Anonymous Repository}

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main 81b14ea82bdf369ba5dbe8de29fa418756900125 \\
       2025-10-15 17:43:32 +0200
}
\titlerunning{WQO on Words}

\author{
    Anonymized for review
}
\authorrunning{Anonymized for review}
\institute{Anonymized for review}

\date{\today}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered. This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science. However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering. In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems. More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions. We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
\end{abstract}

% Include the content of the paper
\input{src/introduction}
\input{src/prelims.tex}
\input{src/prefixes}
\input{src/infixes}
\input{src/amalgam}
\input{src/conclusion}

% Include acknowledgements

% Include the bibliography
\bibliographystyle{splncs04}
\bibliography{papers.bib}

% If there are any appendices, we include them here.
\appendix
\input{src/appendix.tex}

\end{document}