### 2025-12-10 rebuttal.txt
Remove old stuff

We thank the reviewers for their time and their insightful feedback. The
reviewers made relevant comments and suggestions that will be implemented in
the final version of the article.

We first answer the main comments and questions of R1 and R3 and give below a
more detailed response to other comments/questions.

--R1
Comments: We thank R1 for their positive and useful feedback.

Question of R1: Can you expand a bit on the applications you envision your
results can have that cannot be solved with existing tools?

Answer: Almost all wqo results have some form of algorithmic application. For
example, our results can be used as an (effective!) regularity check for trace
languages and safety properties in amalgamation systems.

--R3
Comments: We thank R3 for their appreciation of the quality of the technical level of the paper.
Their main concern is the lack of a single large and complex technical contribution.

Our answer regarding this is twofold: 1) we've put a lot of effort into
breaking down the technical difficulties noticed by the reviewer into
manageable parts, in order to make the proofs as accessible as possible. 2) The
nature of the article is more of a classification paper rather than a solution
to a known hard to solve problem. Indeed, while we are happy that the technical
contributions have been appreciated by the reviewer, we believe that the main
contribution of the article lies in its originality and not in solving a
particularly tough problem.

Question 1: Wouldn't it be more concise to characterise (in Thms 8 and 20) 
IDEALS of languages well quasi ordered by the infix ordering, 
instead of the wqo languages themselves?

Answer: Thank you for this suggestion. Concerning Theorem 8, we think that phrasing it in terms of
ideals would needlessly complicate the statement. For Theorem 20, we are happy
to rephrase it in terms of ideals.

Question 2: How much the characterisation of Lemma 22 can be generalised
beyond well quasi orders, namely for ideals in the infix order?

Answer: The proof can in fact be generalized to all order ideals of the
infix order, and we thank Reviewer 3 for pointing this out. We will adjust
the statement and the proof.

Question 3: Why Corollary 34 does not include VASS languages?
Corollary 34 is meant to emphasize a subpart of our
results that we think might be of interest to a broader audience. The fully
general version of this statement is Theorem 30, which of course includes VASS
and also PVASS languages.

---DETAILED ANSWER TO OTHER COMMENTS

R2 on Corollary 14. We will add a detailed proof in the final version. A sketch
of the argument is that L_k is isomorphic to k copies of N^2, and one can
easily compute its ordinal invariants using standard methods [15].

R2 on Corollary 21, R3 on forward references. We preferred stating early on the
theorems of interest, and then explain which lemmas will be used and how they
fit together in a general proof scheme. This naturally introduces forward
references. As this was found confusing, we will update the paper to be more
linear.

R2 on Lemmata 26 and 24. We will remove the redundant part, thank you for pointing 
it out.

R3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
and we will adjust the statement accordingly.

R3 on Corollary 34. Corollary 34 is meant to emphasize a subpart of our
results that we think might be of interest to a broader audience. The fully
general version of this statement is Theorem 30, which of course includes VASS
and also PVASS languages.

R3 on L.605.  The concatenative property of amalgamation systems simply
requires that each gap word is the concatenation of its constituent gap words
from the smaller runs, but not in which order. We may pick any gap and pick the
ordering in that gap, but the ordering in the other gaps may then be imposed on
us by the structure of the system.

R3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".

R3.
- The product of two languages L and R is the pairwise concatenation of words from L and R.
- "U satisfies P for x, y, and z simultaneously" means that U satisfies P(x), P(y), and P(z).
- "P(n) holds for arbitrarily large values of n" means ∀k ∃n≥k P(n).
- "x≤y or vice versa" means "x≤y or y≤x".