### 2025-12-10 rebuttal.txt
fix: go below 600 words

We thank the reviewers for their time and their insightful feedback. The
reviewers made relevant comments and suggestions that will be implemented in
the final version of the article.

We first answer the main comments and questions of R1 and R3 and give below a
more detailed response to other comments/questions.

--R1
We thank R1 for their positive and useful feedback.

Question of R1: Can you expand a bit on the applications you envision your
results can have that cannot be solved with existing tools?

Answer: Almost all wqo results have some form of algorithmic application. For
example, our results can be used as an (effective!) regularity check for trace
languages and safety properties in amalgamation systems.

--R3
We thank R3 for their appreciation of the quality of the technical level of the paper.
Their main concern is the lack of a single large and complex technical contribution.

Our answer regarding this is twofold: 1) we've put a lot of effort into
breaking down the technical difficulties noticed by the reviewer into
manageable parts, in order to make the proofs as accessible as possible. 2) The
nature of the article is more of a classification paper rather than a solution
to a known hard to solve problem. Indeed, while we are happy that the technical
contributions have been appreciated by the reviewer, we believe that the main
contribution of the article lies in its originality and not in solving a
particularly tough problem.

Answer to Question 1: Thank you for this suggestion. Concerning Theorem 8, we
think that phrasing it in terms of ideals would needlessly complicate the
statement. For Theorem 20, we are happy to rephrase it in terms of ideals.

Answer to Question 2: The proof of Lemma 22 can in fact be generalized to all order ideals of the
infix order, and we thank Reviewer 3 for pointing this out. We will adjust
the statement and the proof.

Answer to Question 3: Corollary 34 is meant to emphasize a subpart of our
results that we think might be of interest to a broader audience. The fully
general version of this statement is Theorem 30, which of course includes VASS
and also PVASS languages.

---minor-comments---

R2 on Corollary 14. We will add a detailed proof in the final version. A sketch
of the argument is that L_k is isomorphic to k copies of N^2, and one can
easily compute its ordinal invariants using standard methods [15].

R2 on Corollary 21, R3 on forward references. We preferred stating early on the
theorems of interest, and then explain which lemmas will be used and how they
fit together in a general proof scheme. This naturally introduces forward
references. As this was found confusing, we will update the paper to be more
linear.

R2 on Lemmata 26 and 24. We will remove the redundant part, thank you for pointing 
it out.

R3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
and we will adjust the statement accordingly.

R3.
- The product of two languages L and R is the pairwise concatenation of words from L and R.
- "U satisfies P for x, y, and z simultaneously" means that U satisfies P(x), P(y), and P(z).
- "P(n) holds for arbitrarily large values of n" means ∀k ∃n≥k P(n).
- "x≤y or vice versa" means "x≤y or y≤x".

----------- MATERIAL USED TO ANSWER THE REVIEWS --------------

Reviewer 1
----------
- Can you expand a bit on the applications you envision your results can have that cannot be solved with existing tools?
Response:

Ours is a theoretical paper, and as such will mostly have theoretical
applications.

Applications of well-quasi-orderings are numerous in verification, ranging from
termination proofs to decidability of inclusion problems. For the sake of the
argument, let us pick one example (among many)  is the use of wqos in so-called
"well-structured transition systems" (WSTS), where the decidability of certain
verification problems (e.g., coverability) is guaranteed by the presence of a
wqo on the state space that is compatible with the transition relation. In some
cases (like lossy channel systems), the subword ordering appears naturally to
order the configurations, but in most cases the natural ordering on words are
the prefix, suffix, and infix orderings.

To apply the WSTS theory in these cases, one way is to ensure that the set of
reachable configurations (which is often a language) is wqo under the relevant
ordering. Our results provide a complete characterization of such languages,
along with effective decision procedures for languages described in a vast
variety of formalisms. This completely solves the question of whether a WSTS
with configurations ordered by prefix, suffix, or infix ordering can be used
for verification. From our characterizations, one can understand that whenever
one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
reachable configurations must be "simple" in a precise sense, which we
interpret as a formal limitation to the expressiveness of such systems.

Reviewer 2
----------
- You should give a proof for the measures of the languages Lk in the proof of Corollary 14.
Response:
We will add a detailed proof in the final version. Here is a sketch of the argument:
The ordinal height of (infinite) languages is always ω,
the ordinal width of a*b* is ω (since it is isomorphic to N^2 with the product order),
hence disjoint copies of a*b* have width ω + .... + ω, hence the width of L_k is ω·k.
Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is ω^2 added k times, hence ω^2 · k.
We are using classical results on ordinal invariants of wqos that we will take from 
[15].

- You could corollary 21 at the end of section 5.1, since its proof call for lemmas 25 and 26. Or move these lemmas to appendix.
Response:
Yes, this was made to improve the flow of the paper by stating important results first,
but in this case it is counterproductive.

- Part of the claim of Lemma 26 seems redundant with Lemma 24.
Response:
Yes, the first part is the same and will be removed.

- Unfortunately space is limited, but it would have been nice to have Figures 6 and 7 in the core of the paper.
Response:
These figures are quite large, and scaling them down would make them less readable.
We will try moving them to the main text by compressing the prefix/suffix section 
that is less central to the main results.

Reviewer 3
----------
- Wouldn't it be more concise to characterise (in Thms 8 and 20) 
IDEALS of languages well quasi ordered by the infix ordering, 
instead of the wqo languages themselves?
Response: For Theorem 8, we think that phrasing it in terms of ideals would needlessly complicate the statement. For Theorem 20, we are happy to rephrase it in terms of ideals.

- How much the characterisation of Lemma 22 can be generalised
beyond well quasi orders, namely for ideals in the infix order?
Response: The proof can in fact be generalized to all order ideals 
of the infix order, and we thank the Reviewer for pointing this out.
We will adjust the statement and the proof.

- Why Corollary 34 does not include VASS languages?
Response: Corollary 34 is meant to emphasize a subpart of our results
that we think might be of interest to a broader audience. The fully general
version of this statement is Theorem 30, which of course includes VASS and 
also PVASS languages.

- There is no a really long and complex proof in the paper, rather a series of
short arguments. As a reader I had an impressions of seeing a series
of small fragmental results, while I'd probably prefer one big result
instead of 20 small ones. 
Response: As the space of infix well-quasi-orders was not well explored so far, we onsider this as a classification paper, which naturally lends itself to a large range of smaller statements rather than singular large ones.

- The section on prefix/suffix could be probably completely removed from the
paper, with the extra space devoted to the most important arguments that are
now in the appendix.
Response: We agree and will condense the section significantly to free up space,
especially to reintroduce the diagrams to the main text as requested by Reviewer 2.

- L.197-199: unclear what is the tree relation here.
Response: The trees are formed as follows: The root node is the empty sequence, ε. Each node has (potentially infinitely many)
children from the wqoset (X, ≤) such that the sequence of elements from the root to this child satisfy the property expressed
by the tree (i.e., being bad, decreasing or an antichain respectively).

- Thm 8: what are "products"?
Response: By the "product" of two languages L and R we mean the usual pairwise concatenation of words from L and R.

- Lemma 10: what does it mean "simultaneously" here?
Response: By "simulatenously" we mean that we can find a finite set of languages such that their union is P↓(x), and each
language is a chain by the prefix relation, and by the infix relation, and by the suffix relation.

- Lemma 12: what is the meaning of "for arbitrarily large values of n" here?
Response: By "for abritrarily large values of n", we formally mean: ∀k ∃n≥k such that the property holds.

- Lemma 12: point 2 seems to be subsumed by 3, isn't it?
Response: Correct, both point 1 and point 2 can be subsumed by point 3. The phrasing was left over from a previous version of the proof. We will adjust this.

- L.605: what it means "we may fix"?
Response: The concatenative property of amalgamation systems simply requires that each gap word is the concatenation of its 
constituent gap words from the smaller runs, but not in which order. We may pick any gap and pick the ordering in that gap,
but the ordering in the other gaps may then be imposed on us by the structure of the system. For example, when amalgamating
derivation trees of CFGs, each gap has a sibling gap on the intuitive "other side" of the tree that must be concatenated in
the reverse order.

- Lemma 32: "or vice versa" not clear.
Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f

- Lemma 32: "or vice versa" not clear.
Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f